Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-05-14 Chapter-5/5-Preparatory.tex
s

\begin{figure}
\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam60A.png}
\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
\caption{Beam composition for the -60A runs (left) and -100A runs (right). The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content. The plots are area normalized to the number of data events, shown in red. }
\label{fig:BeamComposition}
\end{figure}

@@

When producing a DDMC sample, beam line data from a particular running period and/or running condition are selected first. For example, data for the negative 60A runs and for the negative 100A runs inform the event generation stage of two different DDMC samples. Figure \ref{fig:DDMCQuantities}  schematically shows the data quantities of interest leveraged from data: the momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4. For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}. The momentum and position of the selected data

The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousand of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events. This produces MC $P_x, P_y, P_z, X, Y$ distributions  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables. A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the sampled momentum and position distributions as a template. 
As an example, the results of the DDMC generation compared to data for the pion 60A sample are shown in figure \ref{DDMCComparison}; as expected, MC and data agree within the statistical uncertainty by construction. Using this technique ensures the MC and data particles have very similar momentum, position and angular distributions at WC4 and allow us to us the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}) or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}). A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.

Three sample of \textcolor{red}{NUMBERS} pions, muons and electrons, as well as  a sample of \textcolor{red}{NUMBERS} kaons have been generated with the DDMC and are used for the MC cross section study.

@@

\begin{figure}[hpbt]
\centering
\includegraphics[width=\textwidth]{Chapter-5/Images/Pz.png}
\caption{Comparison between generated quantities and data distributions for the 60A pion sample: Z component of the momentum (top left), X position at Wire Chamber 4 (top right), Y position at Wire Chamber 4 (bottom).}
\label{fig:DDMCComparison}
\end{figure}

\subsection{Estimate of Energy Loss before the TPC}\label{ch:eloss}

>>>>

\begin{figure}
\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-5/Images/Beam60A.png}
\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-5/Images/Beam100A.png}
\caption{Beam composition for the -60A runs (left) and -100A runs (right). The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content. The plots are area normalized to the number of data events, shown in red. }
\label{fig:BeamComposition}
\end{figure}

@@

When producing a DDMC sample, beam line data from a particular running period and/or running condition are selected first. For example, data for the negative 60A runs and for the negative 100A runs inform the event generation stage of two different DDMC samples. Figure \ref{fig:DDMCQuantities}  schematically shows the data quantities of interest leveraged from data: the momentum ($P_x, P_y, P_z$) and position ($X, Y$) at WC4. For each data event, we obtain the  particle position ($X, Y$) at WC4 directly from the data measurement; we calculate the components of the momentum using the beamline measurement of the momentum magnitude in conjunction with the hits on WC3 and WC4 to determine the direction of the momentum vector, as described in section \ref{sec:MWPCfunc}. The momentum and position of the selected data form a 5-dimensional tupla, which we sample thousands of times through a 5-dimensional hit-or-miss sampling procedure to generate the MC events. This produces MC $P_x, P_y, P_z, X, Y$ distributions  with the same momentum and position distributions as data, with the additional benefit of accounting for the correlations between the considered variables.  As an example, the results of the DDMC generation compared to data for the kaon +100A sample are shown in figure \ref{DDMCComparison} for the $P_z$, $X$ and $Y$ distributions; as expected, MC and data agree within the statistical uncertainty by construction. A LArSoft simulation module then launches single particle MC from z = -100 cm (the location of the WC4) using the MC generated events. The particles are free to decay and interact in their path from WC4 to the TPC according to the Geant4 simulation.

Using the DDMC technique ensures that the MC and data particles have very similar momentum, position and angular distributions at WC4 and allows us to use the MC sample in several occasions, for example to calibrate the energy loss upstream of the TPC (see Section \ref{ch:eloss}) or to study the tracking and the calorimetric performance (sections \ref{sec:TrackingStudies} and \ref{ch:energyCal}). A small caveat is in order here: the DDMC is a single particle Monte Carlo, which means that the beam pile-up is not simulated.

Six samples are the basis fo the MC used in the pion cross section measurement: three samples of  $\sim$340000 pions, muons and electrons to simulate the negative 60A runs, and three samples of $\sim$340000 pions, muons and electrons for the negative 100A runs.

The MC used for the kaon cross section analysis is a sample of \textcolor{red}{NUMBERS}  kaons.

@@

\begin{figure}[hpbt]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DDMCPz.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DDMCX.png}
\includegraphics[width=0.48\textwidth]{Chapter-5/Images/DDMCY.png}
\caption{Comparison between generated quantities and data distributions for the 100A kaon sample: Z component of the momentum at WC4 (top left), X position at Wire Chamber 4 (top right), Y position at Wire Chamber 4 (bottom).}
\label{fig:DDMCComparison}
\end{figure}

\subsection{Estimate of Energy Loss before the TPC}\label{ch:eloss}
The beamline particles travel a path from where their  momentum is measured in the beamline until they are tracked again inside the TPC. In the LArIAT geometry, a particle leaving the WC4 will encounter the materials listed in Table \ref{tab:budget} before being registered again. The energy lost by the particle in this non-instrumented material modifies the particle's kinetic energy and directly affects the cross section measurement, as shown in equation \ref{eq:enFF}.

\begin{table}[h!]
\centering
\begin{tabular}{|l|l|l|}
\hline
Material  & density {[}g/cm$^3${]} & width {[}cm{]}    \\ \hline
Fiberglass laminate (G10)      & 1.7                             & 1.28                              \\
Liquid Argon                           & 1.4                             & 3.20                             \\
Stainless Steel                        & 7.7                            & 0.23                             \\
Titanium                                  & 4.5                            & 0.04                             \\ 
Air                                            &  1.2 $\cdot10^{-3}$  & 89.43                              \\
Plastic Scintillator                    & 1.03                          & 1.20 (+ 1.30)                             \\ \hline
\end{tabular}
\caption{LArIAT material budget from WC4 to the TPC Front Face.}
\label{tab:budget}
\end{table}

We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Data Driven Monte Carlo using the pion and kaon samples separately, since this quantity is not  measurable directly on data. 
The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively. A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}. The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle. In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation. These distributions can be separated drawing a line in this position-momentum space. 
We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line. We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation. Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
We apply the same classifier on data. We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".

%We use the separation of these two distributions to decide what the energy loss for each event on data. 
%Thus,  we assign the value for energy loss is used in the data.

\begin{figure}[hbpt]
\centering
\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss60A.png}
\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss100A.png}
\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right). The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.  }
\label{fig:ELoss60A}
\end{figure}

\begin{figure}[hbpt]
\centering
\includegraphics[scale=0.5]{Chapter-5/Images/Halo.png}
\caption{Schematic rendering of the particle path between WC4 and the TPC front face. The paddle with the hollow central circle represents the Halo paddle. We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
\label{fig:Halo}
\end{figure}

\begin{figure}[hbpt]
\centering
\includegraphics[width=\textwidth]{Chapter-5/Images/PXVsX60A.png}
\caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs. The plot on the left shows the distribution for pion that miss the halo paddle and the plot on the right shows the distributions for pions that hit the halo. The form of the classifier is overlaid to both plots (red line).}
\label{fig:PxVsXTrue}
\end{figure}

### 2018-05-14 Chapter-8/8-CrossSectionAnalysis.tex
s

\section{Estimate of $E_{loss}$ before the TPC}
The beamline particles travel a path from when their momentum is measured by the beamline detector, until they are tracked again inside the TPC. In the current LArIAT geometry, a particle leaving the fourth wire chamber will encounter the materials listed in Table \ref{tab:budget} before being registered again. The energy lost by the particle in this non-instrumented material modifies the particle's kinetic energy and directly affects the cross section measurement, as shown in equation \ref{eq:enFF}.

\begin{table}[h!]
\centering
\begin{tabular}{|l|l|l|}
\hline
Material  & density {[}g/cm$^3${]} & width {[}cm{]}    \\ \hline
Fiberglass laminate (G10)      & 1.7                             & 1.28                              \\
Liquid Argon                           & 1.4                             & 3.20                             \\
Stainless Steel                        & 7.7                            & 0.23                             \\
Titanium                                  & 4.5                            & 0.04                             \\ 
Air                                            &  1.2 $\cdot10^{-3}$  & 89.43                              \\
Plastic Scintillator                    & 1.03                          & 1.20 (+ 1.30)                             \\ \hline
\end{tabular}
\caption{LArIAT material budget from WC4 to the TPC Front Face.}
\label{tab:budget}
\end{table}

We derive an estimate of the energy loss between the beamline momentum measurement and the TPC ($E_{loss}$) from the Data Driven Monte Carlo using the pion and kaon samples separately, since this quantity is not  measurable directly on data. 
The $E_{loss}$ distribution for the 60A  and 100A pion sample is shown in figure \ref{fig:ELoss60A}, left and right respectively. A clear double peaked structure is visible, which is due to the particles either missing or hitting the HALO paddle: a schematic rendering of this occurrence is  shown in figure \ref{fig:Halo}. The kinematic at WC4 determines the trajectory of a particle and whether or not it will hit the halo paddle. In figure \ref{fig:PxVsXTrue} , we plot the true  $X$ component of the momentum versus the true $X$ position at WC4 for pions missing the halo paddle (left) and for pions hitting the halo paddle (right) for the 60A MC simulation runs -- analogous plots are obtained with the 100A simulation. These distributions can be separated drawing a line in this position-momentum space. 
We use a logistic regression  \cite{agresti2013categorical}  as a classifier to find the best separating line, shown in both plots as the red line. We classify as ``hitting the halo paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02* X - 0.4 < 0 $$ and as ``missing the halo  paddle" all pions whose $P_x$ and $X$ are such that $$P_x +0.02*X - 0.4 > 0, $$ where the coefficients of the line are empirically found by the logistic regression estimation. Overall, this simple classifier classifies in the right category (hit or miss) about 86\% of the pion events.
We apply the same classifier on data. We assign  $E_{loss} = 32 \pm 4 $~MeV for events classified as ``hitting the halo paddle"; we assign  $E_{loss} = 24 \pm 3 $~MeV for events classified as ``missing the halo paddle".

%We use the separation of these two distributions to decide what the energy loss for each event on data. 
%Thus,  we assign the value for energy loss is used in the data.

\begin{figure}[hbpt]
\centering
\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss60A.png}
\includegraphics[width=0.45\textwidth]{Chapter-5/Images/E_loss100A.png}
\caption{True energy loss between WC4 and the TPC front face according to the MC simulation of the 60A runs (left) and of the 100A runs (right). The distribution for the whole data sample is shown in blue, the distribution for the pions missing the halo is shown in red, and the distribution for the pions hitting the halo is shown in green.  }
\label{fig:ELoss60A}
\end{figure}

\begin{figure}[hbpt]
\centering
\includegraphics[scale=0.5]{Chapter-5/Images/Halo.png}
\caption{Schematic rendering of the particle path between WC4 and the TPC front face. The paddle with the hollow central circle represents the Halo paddle. We illustrate two possible trajectories: in black, a trajectory that miss the paddle and goes through the hole in the Halo, in blue a trajectory that hits the Halo paddle and goes through the scintillation material.}
\label{fig:Halo}
\end{figure}

\begin{figure}[hbpt]
\centering
\includegraphics[width=\textwidth]{Chapter-5/Images/PXVsX60A.png}
\caption{Horizontal component of the true momentum vs the horizontal position at WC4 for MC simulated pions of the 60A runs. The plot on the left shows the distribution for pion that miss the halo paddle and the plot on the right shows the distributions for pions that hit the halo. The form of the classifier is overlaid to both plots (red line).}
\label{fig:PxVsXTrue}
\end{figure}

\section{Interacting and Incident Distributions}

>>>>

\section{Estimate of $E_{loss}$ before the TPC}
\section{Interacting and Incident Distributions}
